Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-04-24 WSSSPE1_report/WSSSPE1_summary_report.tex
small fix

A relatively new paradigm in utility software is the software
delivered as service over the web. With increasing popularity of
cloud-based storage and computational environments, many users are
leaning towards tools used as services. \toolname{GitHub} and
\toolname{Bitbucket} can be argued to be such tools, catering to
collaborative development. For scientific users
\toolname{Globus}-based tools are a case of service-based utility
software discussed during the workshop. The data movement and sharing
services offered by Globus can be easily used over the web by
collaborating researchers.
% \note{Dan: so is the only difference here that this category is
%aimed at new systems, where utility software is not?  If so, perhaps
%this should be incorporated into that subsection, so there would just
%be two subsections.  Something in the framing of this subsection
%doesn't feel right to me.  FL: I respectfully disagree the the only
%reason to develop frameworks is to better utilize (new or old)
%infrastructure. Another really important reason is to modularize
%development, making it much easier for an individual developer/user
%to integrate their own code. Ketan: I think in this section I wanted
%to bring up a class of software projects which are undertaken with
%the main aim of tackling new systems and architectures in mind. New
%infrastructures such as GPU, Cloud, fast IB, federal science
%networks, etc. were on my mind as I see many submissions around them
%in the workshop.}

@@

% What are the challenges?
Long-term sustainability of scientific software is often a significant
challenge and face radically different issues compared to utility
software.  Many submissions reported that software is practically
considered a ``byproduct'' of the actual research. Others contended
that the software was not the main funded part of their research. A
smaller codebase and fixed requirements result in stability, ease of
installation, and configurability.  Many such projects mature and are
treated as libraries to be integrated into larger systems such as some of
the utility software discussed in the previous section. While the
software can stay stable and require relatively low maintenance, the
responsibility is often on the shoulders of a few developers who
are often not specialists in software development. Development tends to
be linear and simplistic with a limited scope to follow software best
practices.

>>>>

A relatively new paradigm in utility software is the software
delivered as service over the web. With increasing popularity of
cloud-based storage and computational environments, many users are
leaning towards tools used as services. \toolname{GitHub} and
\toolname{Bitbucket} can be argued to be such tools, catering to
collaborative development. For scientific users
\toolname{Globus}-based tools are a case of service-based utility
software discussed during the workshop. The data movement and sharing
services offered by \toolname{Globus} can be easily used over the web by
collaborating researchers.
% \note{Dan: so is the only difference here that this category is
%aimed at new systems, where utility software is not?  If so, perhaps
%this should be incorporated into that subsection, so there would just
%be two subsections.  Something in the framing of this subsection
%doesn't feel right to me.  FL: I respectfully disagree the the only
%reason to develop frameworks is to better utilize (new or old)
%infrastructure. Another really important reason is to modularize
%development, making it much easier for an individual developer/user
%to integrate their own code. Ketan: I think in this section I wanted
%to bring up a class of software projects which are undertaken with
%the main aim of tackling new systems and architectures in mind. New
%infrastructures such as GPU, Cloud, fast IB, federal science
%networks, etc. were on my mind as I see many submissions around them
%in the workshop.}

@@

% What are the challenges?
Long-term sustainability of scientific software is often a significant
challenge and face radically different issues compared to utility
software.  Many submissions reported that software is practically
considered a ``byproduct'' of the actual research. Others contended
that the software was not the main funded part of their research. A
smaller codebase and fixed requirements result in stability, ease of
installation, and configuration. Many such projects mature and are
treated as libraries to be integrated into larger systems such as some of
the utility software discussed in the previous section. While the
software can stay stable and require relatively low maintenance, the
responsibility is often on the shoulders of a few developers who
are often not specialists in software development. Development tends to
be linear and simplistic with a limited scope to follow software best
practices.
